Case study: Analysis on smart meter data

### 2019-08-25 main.Rmd
figure referencinf correctedremoving special characters

\autoref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year. The two households are similar when it comes to their usage in the morning hours. The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage). However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar. This implies that consuming more energy in the evening is more typical for household 2. For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2. Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH. This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely. All of these can be explained with more demographics of the household and weather conditions in that area.

@@

```{r cust1_hd_my}
library(ggplot2)
data_sm1 %>% 
  granplot("month_year", "hour_day", 
           response = "general_supply_kwh", 
           plot_type = "quantile", 
           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

```{r cust3_hd_my}
data_sm3 %>% 
  create_gran("month_year") %>% 
  filter(month_year != "Mar") %>% 
    granplot("month_year", "hour_day", "general_supply_kwh",
             plot_type = "quantile", 
             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + 
  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

@@

```{r cust1_hd_dw}
data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
```
<!-- # ``````{r cust2_hour-day_day-week} -->
<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
<!-- # ``` -->

``````{r cust3_hd_dw}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
```

\autoref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1. Energy consumption typically goes up between 16 to 18 hours. On weekends, the energy consumption for these hours are even higher. The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour. This suggests that these hours are extremely volatile based on weather conditions/months. Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range. This is not true for the 17th or 18th hour of the day. The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home professionals.

\autoref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2. Contrary to Household 1, Household 2's energy consumption increases through noon through 7pm. We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption from the households. This indicates that this household might have kids or someone working from home.

A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June. Was it only for the months May and June, that this household consumed more energy than they usually do? \autoref{cust1_dw_my} shows that it is not only those months that saw high energy consumption on Saturday and Sundays. In fact, months like November and December saw higher energy usage than typical even on other weekdays. Still for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends. The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.

/autoref{fig:cust3_dw_my} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week. The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December. Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.

@@

```{r cust1_dw_my}
library(ggplot2)

data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv", quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_dw_my}
library(ggplot2)

data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 3")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

@@

 - First household tpically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June). They are strictly 9 to 5 professsionals, hence their energy consumption during day time is pretty consistent and only peaks up during few hours in the evening. They do all energy intensive chores (e.g. laundry, vacuum cleaning) mostly in the evenings of weekends.

@@

>>>>

Figure \ref{fig:cust1_hd_my} and \ref{fig:cust3_hd_my} shows the distribution of hourly consumption of household 1 and 2 across different months of the year. The two households are similar when it comes to their usage in the morning hours. The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage). However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar. This implies that consuming more energy in the evening is more typical for household 2. For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2. Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH. This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely. All of these can be explained with more demographics of the household and weather conditions in that area.

@@

```{r cust1_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the first household. 10th, 25th, 50th, 75th and 90th percentile are shown. For example, 90th percentile might denote behavior which are extreme and 50th percentile corresponds to a more typical behavior. Evening hours generally exhibits a peak. Some extreme behavior is exhibited in the evening hours of May and June, however, overall energy comsumption does not seem to vary much across months. Median and 75th percentile is less than 0.25 KwH across most months."}
library(ggplot2)
data_sm1 %>% 
  granplot("month_year", "hour_day", 
           response = "general_supply_kwh", 
           plot_type = "quantile", 
           quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) +
  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

```{r cust3_hd_my, fig.cap="Quantile plots of hourly consumption across different months for the second household. 10th, 25th, 50th, 75th and 90th percentile are shown. Distinct evening peaks for few hours across all months. The median comsumption also varies across months. This household varies their behavior with seasons and are likely to be using heaters in colder months and air conditioners in hot months like January. This is evident from the 75th and 90th percentile of consumption."}
data_sm3 %>% 
  create_gran("month_year") %>% 
  filter(month_year != "Mar") %>% 
    granplot("month_year", "hour_day", "general_supply_kwh",
             plot_type = "quantile", 
             quantile_prob = c(0.1, 0.25, 0.5, 0.75, 0.9)) + 
  ggtitle("Quantile plot of energy consumption across hours of day by month of years for Household 2") + 
  scale_y_continuous(breaks = seq(0, 1.5, 0.25)) +
  scale_x_discrete(breaks = seq(0,23,2))
```

@@

Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for household 1. Energy consumption typically goes up between 16 to 18 hours. On weekends, the energy consumption for these hours are even higher. The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour. This suggests that these hours are extremely volatile based on weather conditions/months. Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range. This is not true for the 17th or 18th hour of the day. The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home parents/ working from home professionals.

Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for Household 2. Contrary to household 1, household 2's energy consumption increases through noon to 7pm. We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption for this household. This indicates that this household might have kids or someone working from home.

```{r cust1_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the first household.  A long tailed violin for an hour would imply that the energy distribution is skewed to either right or left, whereas a bulge shaped distribution implies that the energy behavior is more predictable and energy consumption in those hours can only vary in that range where the bulge appears. If an hour has a distribution that seems like a straight line, they are extremely volatile. On weekdays, hours between 17 to 19 hours seems to be extremely volatile comapred to other hours. On weekends, energy consumption is volatile even in morning hours. "}

data_sm1 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") +  ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 1") + scale_x_discrete(breaks = seq(0,23,3))
```
<!-- # ``````{r cust2_hour-day_day-week} -->
<!-- # data_sm2 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "quantile") + ggtitle("Boxplot of energy consumption for Customer 2") -->
<!-- # ``` -->

```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
```

@@

A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June. Was it only for the months May and June, that this household consumed more energy than they usually do \? Figure \ref{fig:cust1_my_dw} shows that it is not only those months that saw high energy consumption on Saturday and Sundays. In fact, months like November and December saw higher energy usage than typical even on other weekdays. Still for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends. The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.

Figure \ref{fig:cust3_my_dw} shows that the second household uniformly distribute their regular activity like using dish washer or heater/air conditioner throughout the week. The distribution of energy consumption on Monday, Friday and Sunday looks quite similar for December. Also, for other months, it is not very clear if weekend behavior varies distinctly from weekday behavior.

@@

 - First household tpically doesn't routinely use air conditioners in summer(December), or heaters in winter months (May and June). They are strictly 9 to 5 professsionals, hence their energy consumption during day time is pretty consistent and only peaks up during few hours in the evening. They do all energy intensive chores (e.g. laundry, vacuum cleaning) mostly in the weekends.

@@

```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations. M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.Outliers are colored in red. It can be noted that they do not typically consume more than 0.25 KwH in the weekdays. Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}

data_sm1 %>% granplot("month_year", 
                      "day_week", 
                      response = "general_supply_kwh", plot_type = "lv") +    ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH. The extreme values in the tail of the distribution are represented by different letter values. In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December. For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}

data_sm3 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days of week by month of years for Household 2")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

### 2019-08-25 smart_meter.Rmd
figure referencinf correctedremoving special characters

Figure \ref{fig:cust1_hd_my} shows the distribution of hourly consumption of the households across different months of the year. The two households are similar when it comes to their usage in the morning hours. The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage). However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar. This implies that consuming more energy in the evening is more typical for household 2. For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2. Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH. This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely. All of these can be explained with more demographics of the household and weather conditions in that area.

@@

(ref:cust1_hd_my-cap) Something that I want to describe.

@@

Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 1. Energy consumption typically goes up between 16 to 18 hours. On weekends, the energy consumption for these hours are even higher. The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour. This suggests that these hours are extremely volatile based on weather conditions/months. Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range. This is not true for the 17th or 18th hour of the day. The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home professionals.

Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage for two households across different days of the week through violin plots for Household 2. Contrary to Household 1, Household 2's energy consumption increases through noon through 7pm. We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption from the households. This indicates that this household might have kids or someone working from home.

@@

```{r cust3_hd_dw, fig.cap="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
```

@@

A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June. Was it only for the months May and June, that this household consumed more energy than they usually do? Figure \ref{fig:cust1_my_dw} shows that it is not only those months that saw high energy consumption on Saturday and Sundays. In fact, months like November and December saw higher energy usage than typical even on other weekdays. Still for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends. The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.

@@

```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations. M, F, E, D and C represents 50%, 25%, 12.5%, 6.25% and 3.13% of the tail area respectively.Outliers are colored in red. It can be noted that they do not typically consume more than 0.25 KwH in the weekdays. Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}

data_sm1 %>% granplot("month_year", "day_week", response = "general_supply_kwh", plot_type = "lv") +  ggtitle("LV plot of energy consumption across days-week by month-year for Household 1")+ scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH. The extreme values in the tail of the distribution are represented by different letter values. In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays (represented by letter value D), which is far more higher than subsequent months and comparable to summer month December. For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}

>>>>

Figure \ref{fig:cust1_hd_my} and \ref{fig:cust3_hd_my} shows the distribution of hourly consumption of household 1 and 2 across different months of the year. The two households are similar when it comes to their usage in the morning hours. The first household's extreme behavior is different from that of second household in the month of May and June in the evening hours (represented by higher 90th percentile energy usage). However, the 75th percentile of energy consumption for those two months are higher for the second household than for the first and their median consumption is more or less similar. This implies that consuming more energy in the evening is more typical for household 2. For the evenings of October, November and December, it is clear that household 1 typically consumes much less energy than household 2. Moreover, the second household has a steep evening peak (75th percentile above 0.50 KwH) across most months in contrast to the first household who has a steep evening peak only in the month of May and June with 75th percentile consumption falling below 0.25 KwH. This might be an indication that of usage of heater in the winter months for the first household, whereas the second household uses both air conditioner in hot months and heater in cold months or the number of members in the household vary widely. All of these can be explained with more demographics of the household and weather conditions in that area.

@@

@@

Figure \ref{fig:cust1_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for household 1. Energy consumption typically goes up between 16 to 18 hours. On weekends, the energy consumption for these hours are even higher. The straight lines in the violin plot for weekends at 17th and 18th hours suggest that the distribution of energy consumption is uniform demonstrating that any energy value has equal probability in falling in that hour. This suggests that these hours are extremely volatile based on weather conditions/months. Energy consumption distribution in the morning hours of weekdays have a bulge between 0 to 0.25 KwH, indicating that if we pick up any typical week day, we can say with reasoable certainty that the energy consumption would fall in that range. This is not true for the 17th or 18th hour of the day. The consistent behavior in the morning hours can be attributed to the fact that members in this household are not stay-at-home parents/ working from home professionals.

Figure \ref{fig:cust3_hd_dw} shows distribution of hourly energy usage across different days of the week through violin plots for Household 2. Contrary to household 1, household 2's energy consumption increases through noon to 7pm. We can't say with certainty if the morning hours from 10am would manifest a higher or lower level of energy consumption for this household. This indicates that this household might have kids or someone working from home.

@@

```{r cust3_hd_dw, fig.cap ="Violin plots of hourly consumption across different day of the week for the second household. Hours before 10am and after 10pm have a bulge implying those hours are pretty consistent. Hours in between these have a long right tail which are even higher for weekends."}
data_sm3 %>% granplot("day_week", "hour_day", "general_supply_kwh", plot_type = "violin") + ggtitle("Violin plot of energy consumption across hour-of-day and day-of-week for Household 2") + scale_x_discrete(breaks = seq(0,23,3))
```

@@

A plot of the distribution across months of the year and days of the week might throw some light on the extreme behavior we saw in May and June. Was it only for the months May and June, that this household consumed more energy than they usually do \? Figure \ref{fig:cust1_my_dw} shows that it is not only those months that saw high energy consumption on Saturday and Sundays. In fact, months like November and December saw higher energy usage than typical even on other weekdays. Still for this household, it can be safely said that most of their activities which involve higher energy consumption are concentrated on weekends. The behavior of the weekends, as can be observed from the letter values of energy consumption, are different across months which can be attributed to weather conditions.

@@

```{r cust1_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Letter value plots convey detailed information about tails of the distribution and outliers are unexpected observations rather than extreme observations. M, F, E, D and C represents 50, 25, 12.5, 6.25 and 3.13 percent of the tail area respectively.Outliers are colored in red. It can be noted that they do not typically consume more than 0.25 KwH in the weekdays. Their tail behaviors are not so distinctly visible because of the consistency in their behavior."}

data_sm1 %>% granplot("month_year", 
                      "day_week", 
                      response = "general_supply_kwh", plot_type = "lv") +    ggtitle("LV plot of energy consumption across days-week by month-year for Household 1") +
  scale_y_continuous(breaks = seq(0, 1.5, 0.25))
```

```{r cust3_my_dw, fig.cap = "Letter value plots of energy consumption across days of the week and months for the first household. Contrary to the first household, outliers appear at an energy usage beyond 0.75 KwH. The extreme values in the tail of the distribution are represented by different letter values. In March, once can observe the F letter value to range between 0.75 to 1.25 KwH on weekends and between 0.25 to 0.75 KwH on weekdays -represented by letter value D, which is far more higher than subsequent months and comparable to summer month December. For unusual days or more volatile days of the week, one can expect to see more distinct letter values.This reinforces the fact that their behavior changes across months but not so much between days of the week."}
